---
title: "HighVariationFST"
author: "Georgina Robertson"
date: "11/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Here the FST output from vcftools FST will be used
________________________________________________________
1a) Preparing the file for first time use
```{r}
fst_file <- read.csv("./data_sheets/pop1_vs_pop2.weir.fst", header = T, sep = "\t")
```

There are some negative vales here. These occur due to the incluson of standarrd deviation in the caluculations. They should be treated as zero, so for clarity they will be changed.
```{r}
# check that the classes are correct
str(fst_file)

# change negative values to zero
fst_file$WEIR_AND_COCKERHAM_FST[fst_file$WEIR_AND_COCKERHAM_FST<0] <- 0
```

```{r}
colnames(fst_file) <- c("Scaffold", "Position", "FST")

Chrom <- sapply(strsplit(fst_file$Scaffold, "_", "_") ,"[", 3)

Chrom <- as.numeric(Chrom)

fst_file <- cbind(fst_file, Chrom)
fst_file <- fst_file[, c(1, 4,2,3)]

write.csv(fst_file, "./data_sheets/fst_file.csv")
```

_________________________________________________________

1b) Reading in formatted file (future use)
```{r}
fst_file <- read.csv("./data_sheets/fst_file.csv")
fst_file <- fst_file[,-1]
```

1c) Reading in file that contains SNP index and FST data which was made in the pipeline: "SnpIndex_FST.Rmd"
```{r}
FST_SnpIndex <- read.csv("./data_sheets/FST_SNPindex.csv")
FST_SnpIndex <- FST_SnpIndex[,-1]
```
_________________________________________________________

2) Manhattan plot
```{r}
library(qqman)

manhattan(fst_file, chr = "Chrom", bp = "Position", p = "FST", logp = F, snp = "Scaffold")

# Saved in /IMGS as Manhattan_All_FST.png
```


3) Now I will look at strong differences between our populations by subsetting loci with <= 0.9 fst

Will use the file from 1c) so I also have the SNP index for corss-referencing the heatmaps

```{r}
ninetyplus <- subset(FST_SnpIndex, FST_SnpIndex$FST >= 0.9)
```

```{r}
library(ggplot2)
```

4.1) Chromosome 1
```{r}
chr1high <- ninetyplus[ninetyplus$Chrom==1, ]
ggplot()+
geom_point(data = chr1high, aes(x=Position, y = FST), colour = "red") +
  xlab("loci")+
  ylab("fst")
# not saved thsi image, only showed 3 points
```

Not many in chromosome 1
This may change if i reduced the threshold to >0.8
```{r}
eightyplus <- subset(FST_SnpIndex, FST_SnpIndex$FST >= 0.8)
```

```{r}
chr1high <- eightyplus[eightyplus$Chrom==1, ]
ggplot()+
geom_point(data = chr1high, aes(x=Position, y = FST), colour = "red") +
  xlab("loci")+
  ylab("fst")

# not saved this image, only showed 4 points
```
Still not many, but does show one loci of high variation
```{r}
print(chr1high$SNPindex)
```
Check heatmap @ 24772 24773 24774 50279

4.2) No  high variation in chrs 2-4
note: can check "eightyplus" later for chr 2,3,4

4.3) A lot of high FST values in chromosome 5
```{r}
chr5high <- ninetyplus[ninetyplus$Chrom==5,]

ggplot()+
geom_point(data = chr5high, aes(x=Position, y = FST), colour = "red") +
  xlab("loci")+
  ylab("fst")+
  ggtitle("High variance SNPs between populations in chromosome 5")+
   theme(plot.title = element_text(hjust = 0.5))
# saved in /IMGS as chr5_HighVar.png
```
Formatting  a variable out of the data to enable easy enembl genome annotation search 
```{r}
# determine the loci range of interest
lower <- head(chr5high$Position, n=1)
upper <- tail(chr5high$Position, n=1)

# this puts the info into a vector that can be copy pasted to ensemble
chr5LociOfInterest <- paste(chr5high$Chrom[[1]],":", lower, "-", upper, sep = "")
```

Getting SNP indexes.
There are 1124 observations, so will store as a vector rather than print it out
```{r}
chr5HighVar <- chr5high$SNPindex
range(chr5HighVar)
```
In the heatmap, this corresponds to the region 16560 to 18211

Which other chromosomes have regions of interest?
```{r}
print(unique(ninetyplus$Chrom))
```
4.4) Looking at chrom 6 next:
```{r}
chr6high <- ninetyplus[ninetyplus$Chrom==6,]

ggplot()+
geom_point(data = chr6high, aes(x=Position, y = FST), colour = "blue") +
  xlab("loci")+
  ylab("fst")+
  ggtitle("High variance SNPs between populations in chromosome 6")+
   theme(plot.title = element_text(hjust = 0.5))
# saved in /IMGS as chr6_HighVar.png
```
In contrast to chromosome 5, this show 2 (maybe 3) distinct regions. *This may change if i lowered the threshhold for the fst value to =>0.8 rather than =>0.9*

Here i will repeat the workflow for producing snp indexes and coordinates for using in further investigations.
```{r}
# determine the loci range of interest
lower <- head(chr6high$Position, n=1)
upper <- tail(chr6high$Position, n=1)

# this puts the info into a vector that can be copy pasted to ensemble
chr6LociOfInterest <- paste(chr6high$Chrom[[1]],":", lower, "-", upper, sep = "")

```

There are less observations in this set, and 2 or 3 distinct regions, so I will print the SNP indexes and eyeball it
```{r}
chr6HighVar <- chr6high$SNPindex
print(chr6HighVar)
```
So the first peak is ~ 13800
Second peak ~ 18000 to 19100
Third peak (pretty tightly grouped!) ~ 20100 to 20980


```{r}
print(unique(ninetyplus$Chrom))
```
4.5) Now for chr7
```{r}
chr7high <- ninetyplus[ninetyplus$Chrom==7,]

ggplot()+
geom_point(data = chr7high, aes(x=Position, y = FST), colour = "darkgreen") +
  xlab("loci")+
  ylab("fst")+
  ggtitle("High variance SNPs between populations in chromosome 7")+
   theme(plot.title = element_text(hjust = 0.5))
# saved in /IMGS as chr7_HighVar.png
```
This is more diffuse than chr6

```{r}
# determine the loci range of interest
lower <- head(chr7high$Position, n=1)
upper <- tail(chr7high$Position, n=1)

# this puts the info into a vector that can be copy pasted to ensemble
chr7LociOfInterest <- paste(chr7high$Chrom[[1]],":", lower, "-", upper, sep = "")

rm(lower, upper)
```

As this was less defined, i will just take a the range of snp indexes
```{r}
chr7HighVar <- chr7high$SNPindex
range(chr7HighVar)
```
[1] 14707 19454

quick reminder for me:
```{r}
print(unique(ninetyplus$Chrom))
```

4.6) Now for chr9
```{r}
chr9high <- ninetyplus[ninetyplus$Chrom==9,]

ggplot()+
geom_point(data = chr9high, aes(x=Position, y = FST), colour = "purple") +
  xlab("loci")+
  ylab("fst")
```
This is only one snp
Maybe a reduction in fst value threshold would he helpful here.
```{r}
chr9high <- eightyplus[eightyplus$Chrom==9,]
ggplot()+
geom_point(data = chr9high, aes(x=Position, y = FST), colour = "purple") +
  xlab("loci")+
  ylab("fst")
```
With changed threshhold (>0.8) There are still few snps of high variation.
Note: Not saying that this is irrelevant, just that it will be difficult to visualise on the heatmap, so will only extract the loci and not bother with getting the snp index.
```{r}
print(chr9high$Position)
```
[1] 2132887 2132890


4.7) Chromosome 15
Only 2 SNPS for >0.9, so will try with >0.8
```{r}
chr15high <- eightyplus[eightyplus$Chrom==15,]

ggplot()+
geom_point(data = chr15high, aes(x=Position, y = FST), colour = "purple") +
  xlab("loci")+
  ylab("fst")
# image not saved
```
Again, low amount of data here, and also is an accessory chromosome which are not considered essential-to-survival

4.8) Finally, chromosome 16, another accessory chromosome
```{r}
chr16high <- ninetyplus[ninetyplus$Chrom==16,]

ggplot()+
geom_point(data = chr16high, aes(x=Position, y = FST), colour = "purple") +
  xlab("loci")+
  ylab("fst")+
  ggtitle("High variance SNPs between populations in chromosome 16")+
   theme(plot.title = element_text(hjust = 0.5))
# saved in /IMGS as chr16_HighVar.png
```

```{r}
# determine the loci range of interest
lower <- head(chr16high$Position, n=1)
upper <- tail(chr16high$Position, n=1)

# this puts the info into a vector that can be copy pasted to ensemble
chr16LociOfInterest <- paste(chr16high$Chrom[[1]],":", lower, "-", upper, sep = "")
```

```{r}
chr16HighVar <- chr16high$SNPindex
print(chr16HighVar)
```
First peak spans ~ 1365 to 1547
Second ~ 1971 to 1982
Third, just one blip @ 2839

5) Build a small file with the loci of interest for checking ensembl

```{r}
write(c(chr5LociOfInterest, chr6LociOfInterest, chr7LociOfInterest, chr16LociOfInterest), file = "./data_sheets/highvariation_byLOCI.txt")
```
