---
title: "binning_LD"
author: "Georgina Robertson"
date: "22/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Visualising linkage decay

Thanks to this amazing blog post: https://www.jdatalab.com/data_science_and_data_mining/2017/01/30/data-binning-plot.html

Reading in and formatting of the data
```{r}
chr13postLD <- read.delim("./LDfiles/post_chr13.ld", header = F)
# peek
head(chr13postLD)
# col names
colnames(chr13postLD) <- c("CHR",	"POS1", "POS2",	"N_CHR",	"Rsq",	"D",	"Dprime")
# new varible of distance
chr13postLD$distance <- chr13postLD$POS2 - chr13postLD$POS1
```
__________________________________________________________________________________________

1) Count of binned distances

Define bins
```{r}
breaks <- c(0,1000,  2000, 3000, 4000 ,5000, 6000, 7000 , 8000 , 9000, 10000, 11000,
            12000, 13000, 14000, 15000, 16000, 17000, 18000, 19000, 20000, 21000, 22000,
            23000, 24000, 25000)
```

Define bin labels
```{r}
tags <- c("[0-1)","[1-2)", "[2-3)", "[3-4)", "[4-5)", "[5-6)",
          "[6-7)", "[7-8)","[8-9)", "[9-10)", "[10-11)","[11-12)",
          "[12-13)", "[13-14)", "[14-15)", "[15-16)","[16-17)", 
          "[17-18)","[18-19)", "[19-20)","[20-21)","[21-22)",
          "[22-23)", "[23-24)", "[24-25)")
```

Add the data to the bins
```{r}
group_tags <- cut(v$distance, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
```

Check it out
```{r}
summary(group_tags)
```

```{r}
library(ggplot2)
library(dplyr)
```
Plot couunt of binned distances
```{r}
ggplot(data = as_tibble(group_tags), mapping = aes(x=value)) + 
  geom_bar(fill="burlywood3",color="white",alpha=0.7) + 
  labs(x='mean distance(kbp)') +
  theme_minimal()
```
______________________________________________________________________________

2) Plotting average Rsq per bin

Subset main data
```{r}
v <- chr13postLD %>% select(Rsq, distance)
```

Assign bin labels
```{r}
tags <- c("[0-1)", "[1-2)", "[2-3)", "[3-4)", "[4-5)", "[5-6)", "[6-7)", "[7-8)", "[8-9)", "[9-10)", "[10-11)", "[11-12)", "[12-13)", "[13-14)", "[14-15)", "[15-16)", "[16-17)", "[17-18)", "[18-19)", "[19-20)", "[20-21)", "[21-22)", "[22-23)", "[23-24)", "[24-25)")
```

Bin distances, store groups as new column
```{r}
vgroup <- as_tibble(v) %>% 
  mutate(tag = case_when(
    distance < 1000 ~ tags[1],
    distance >= 1000 & distance < 2000 ~ tags[2],
    distance >= 2000 & distance < 3000 ~ tags[3],
    distance >= 3000 & distance < 4000 ~ tags[4],
    distance >= 4000 & distance < 5000 ~ tags[5],
    distance >= 5000 & distance < 6000 ~ tags[6],
    distance >= 6000 & distance < 7000 ~ tags[7],
    distance >= 7000 & distance < 8000 ~ tags[8],
    distance >= 8000 & distance < 9000 ~ tags[9],
    distance >= 9000 & distance < 10000 ~ tags[10],
    distance >= 10000 & distance < 11000 ~ tags[11],
    distance >= 11000 & distance < 12000 ~ tags[12],
    distance >= 12000 & distance < 13000 ~ tags[13],
    distance >= 13000 & distance < 14000 ~ tags[14],
    distance >= 14000 & distance < 15000 ~ tags[15],
    distance >= 15000 & distance < 16000 ~ tags[16],
    distance >= 16000 & distance < 17000 ~ tags[17],
    distance >= 17000 & distance < 18000 ~ tags[18],
    distance >= 18000 & distance < 19000 ~ tags[19],
    distance >= 19000 & distance < 20000 ~ tags[20],
    distance >= 20000 & distance < 21000 ~ tags[21],
    distance >= 21000 & distance < 22000 ~ tags[22],
    distance >= 22000 & distance < 23000 ~ tags[23],
    distance >= 23000 & distance < 24000 ~ tags[24],
    distance >= 24000 & distance < 25000 ~ tags[25]
   ))
summary(vgroup)
```
Make tag a factor variable for potting
```{r}
vgroup$tag <- factor(vgroup$tag,
                       levels = tags,
                       ordered = FALSE)
summary(vgroup$tag)
```
Boxplot of avg LD per bin
```{r}
ggplot(data = vgroup, mapping = aes(x=tag,y=Rsq)) + 
  geom_boxplot(fill="burlywood3",color="darkslategrey",alpha=0.3) + 
  # geom_jitter(aes(color='blue'),alpha=0.2) +
  labs(x="distance (kbp)") +
  ggtitle("Average LD decay of post-drought population in chromosome 13") +
  theme(plot.title=element_text(hjust = 0.5)) +
  guides(color=FALSE)

```


> Images saved in /IMGS/ as:
avg_LDdecay_POST|PRE_Chr#.png






