---
title: "PlotFilteredSNPS"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## First replace FILTERED1,FILTERED5 with simple FILTER in .tab file

 # Import VCF files
```{r}
getwd()
UNFILT<-read.delim("/Volumes/Solomon_Lab/Megan/ZYMO_GWAS/June2019_rajinMappedReads/GATK_vcfs/joint_genotype/AUS_NZ_450.plain.first.tab", header=T, stringsAsFactors=T)

head(UNFILT)
library("ggplot2")
```
## 1. Straight Quality plot
```{r}
library("ggplot2")
UNQUALplot<-ggplot(UNFILT, aes(x=QUAL))+geom_histogram(alpha=0.5)+ xlim(0,2000)
UNQUALplot

QUALplot<-ggplot(UNFILT, aes(x=log(QUAL)))+geom_histogram(aes(fill=FILTER), alpha=0.5)+
  theme(legend.position = "none")
QUALplot

```


## 2. Quality by Depth Plot (QD)
```{r, echo=FALSE}
UNQDplot<-ggplot(UNFILT, aes(x=QD))+geom_histogram()
UNQDplot

QDplot<-ggplot(UNFILT, aes(x=QD))+geom_histogram(aes(fill=FILTER), alpha=0.5)+
  theme(legend.position = "none")
QDplot

```

## 3. FisherStrand plot (FS)
some values at 0 dont work for the log10 scale so I added a very small amount to the variable (0.1)
```{r}
UNFSplot<-ggplot(UNFILT, aes(x=FS))+geom_histogram()
UNFSplot

FSplot<-ggplot(UNFILT, aes(x=FS))+geom_histogram(aes(fill=FILTER), alpha=0.5)+xlim(-1,25)+
  theme(legend.position = "none")
FSplot


```

## 4. StrandOddsRatio (SOR)
```{r}
UNSORplot<-ggplot(UNFILT, aes(x=SOR))+geom_histogram()
UNSORplot

SORplot<-ggplot(UNFILT, aes(x=SOR))+geom_histogram(aes(fill=FILTER), alpha=0.5)+
  theme(legend.position = "none")
SORplot

```

## 5. RMSMappingQuality (MQ)
```{r}
UNRMSplot<-ggplot(UNFILT, aes(x=MQ))+geom_histogram()
UNRMSplot

RMSplot<-ggplot(UNFILT, aes(x=MQ))+geom_histogram(aes(fill=FILTER), alpha=0.5)+
  theme(legend.position = "none")
RMSplot

```

## 6. MapplingQualityRankSumTest (MQRankSum)
```{r}
UNMQrankplot<-ggplot(UNFILT, aes(x=MQRankSum))+geom_histogram()
UNMQrankplot

MQrankplot<-ggplot(UNFILT, aes(x=MQRankSum))+geom_histogram(aes(fill=FILTER), alpha=0.5)+
  theme(legend.position = "none")
MQrankplot

```
## 7. ReadPosRankSumTest (ReadPosRankSum)
```{r}
UNRPrank<-ggplot(UNFILT, aes(x=ReadPosRankSum))+geom_histogram()
UNRPrank

RPrank<-ggplot(UNFILT, aes(x=ReadPosRankSum))+geom_histogram(aes(fill=FILTER), alpha=0.5)+
  theme(legend.position = "none")
RPrank

```

### Get Legend for shared plotting
```{r}
library(cowplot)
legendp<-ggplot(UNFILT, aes(x=ReadPosRankSum))+geom_histogram(aes(fill=FILTER), alpha=0.5)

legend <- get_legend(
  # create some space to the left of the legend
  legendp + theme(legend.box.margin = margin(0, 0, 0, 12))
)
legend
```

## All 
```{r}


Combined<-plot_grid(QUALplot,QDplot,FSplot,SORplot,RMSplot,MQrankplot,RPrank, align="v", ncol = 2, axis="lrtb") 
Combined
getwd()
ggsave("/Volumes/Solomon_Lab/Megan/ZYMO_GWAS/June2019_rajinMappedReads/GATK_vcfs/joint_genotype/NZ_AUS450_qualFilters.pdf", Combined, width = 24, height =32, units=c("cm"),dpi=300)


```

