---
title: "Pop_Size_Bias_Check"
author: "G.Robertson"
date: "14/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Population size bias 

https://scientificallysound.org/2018/04/19/small-sample-sizes-and-the-bias-of-small-numbers/


Problem:
Of our two populations, the post drought population is > 10x the size of the pre drough population. This may have affected the results.

Checking this:
Create some smaller vcf files from post population with n=27

Getting post isolate list and selecting 5 random picks of 27 isolates
```{r}
postpop <- read.delim(file = "./data_sheets/postDsamples.txt", header = F)

s1 <- sample(postpop$V1, 29, replace = F)
write(s1, file = "s1.keep")
s2 <- sample(postpop$V1, 29, replace = F)
write(s2, file = "s2.keep")
s3 <- sample(postpop$V1, 29, replace = F)
write(s3, file = "s3.keep")
s4 <- sample(postpop$V1, 29, replace = F)
write(s4, file = "s4.keep")
s5 <- sample(postpop$V1, 29, replace = F)
write(s5, file = "s5.keep")
```

These files I will move to the direcctory that contains the post pop vcf.
Then perform:
```{bash}
#1
STARTING_VCF="post_pops.recode.vcf"
KEEP="s1.keep"
OUTVCF="post.s1"

vcftools --vcf ${STARTING_VCF} --keep ${KEEP} --recode --out ${OUTVCF}

#2
STARTING_VCF="post_pops.recode.vcf"
KEEP="s2.keep"
OUTVCF="post.s2"

vcftools --vcf ${STARTING_VCF} --keep ${KEEP} --recode --out ${OUTVCF}

#3
STARTING_VCF="post_pops.recode.vcf"
KEEP="s3.keep"
OUTVCF="post.s3"

vcftools --vcf ${STARTING_VCF} --keep ${KEEP} --recode --out ${OUTVCF}

#4
STARTING_VCF="post_pops.recode.vcf"
KEEP="s4.keep"
OUTVCF="post.s4"

vcftools --vcf ${STARTING_VCF} --keep ${KEEP} --recode --out ${OUTVCF}

#5
STARTING_VCF="post_pops.recode.vcf"
KEEP="s5.keep"
OUTVCF="post.s5"

vcftools --vcf ${STARTING_VCF} --keep ${KEEP} --recode --out ${OUTVCF}
```

Output vcfs and log files moved to new folder /Bias_Check


Lets do Ndiv and TajD first as they are fairly easy to compute and summarise.
Hopefully the graphs will look similar to the ones we already have
Nucleotide diversity calculation
```{bash}
#1
INPUT_VCF="post.s1.recode.vcf"
OUTFILE="ND_s1_post"

vcftools --vcf ${INPUT_VCF} --haploid --window-pi  10000 --out ${OUTFILE}

#2
INPUT_VCF="post.s2.recode.vcf"
OUTFILE="ND_s2_post"

vcftools --vcf ${INPUT_VCF} --haploid --window-pi  10000 --out ${OUTFILE}

#3
INPUT_VCF="post.s3.recode.vcf"
OUTFILE="ND_s3_post"

vcftools --vcf ${INPUT_VCF} --haploid --window-pi  10000 --out ${OUTFILE}

#4
INPUT_VCF="post.s4.recode.vcf"
OUTFILE="ND_s4_post"

vcftools --vcf ${INPUT_VCF} --haploid --window-pi  10000 --out ${OUTFILE}

#5
INPUT_VCF="post.s5.recode.vcf"
OUTFILE="ND_s5_post"

vcftools --vcf ${INPUT_VCF} --haploid --window-pi  10000 --out ${OUTFILE}
```

Read them into R
```{r}
pi1 <- read.delim("./Bias_Check/ND_s1_post.windowed.pi", header = T)
pi2 <- read.delim("./Bias_Check/ND_s2_post.windowed.pi", header = T)
pi3 <- read.delim("./Bias_Check/ND_s3_post.windowed.pi", header = T)
pi4 <- read.delim("./Bias_Check/ND_s4_post.windowed.pi", header = T)
pi5 <- read.delim("./Bias_Check/ND_s5_post.windowed.pi", header = T)
```

Plots

```{r}
library(ggplot2)
```
```{r}
pi_pl1 <- 
  ggplot(pi1, aes(x=PI))+
  geom_histogram(bins = 20, fill="darkslategrey", color="burlywood3")+
  ylab("nDiv")+
  xlab("")+
ggtitle("Sample 1")

pi_pl2 <- 
   ggplot(pi2, aes(x=PI))+
  geom_histogram(bins = 20, fill="darkslategrey", color="burlywood3")+
  ylab("nDiv")+
     xlab("")+
ggtitle("Sample 2")
 
pi_pl3 <- 
   ggplot(pi3, aes(x=PI))+
  geom_histogram(bins = 20, fill="darkslategrey", color="burlywood3")+
  ylab("nDiv")+
     xlab("")+
ggtitle("Sample 3")
 
pi_pl4 <- 
   ggplot(pi4, aes(x=PI))+
  geom_histogram(bins = 20, fill="darkslategrey", color="burlywood3")+
  ylab("nDiv")+
     xlab("")+
ggtitle("Sample 4")
 
pi_pl5 <- 
   ggplot(pi4, aes(x=PI))+
  geom_histogram(bins = 20, fill="darkslategrey", color="burlywood3")+
  ylab("nDiv")+
     xlab("")+
ggtitle("Sample 5")
```

Now for Taj D
```{bash}
#1
INPUT_VCF="post.s1.recode.vcf"
OUTFILE="TJ_s1_post"

vcftools --vcf ${INPUT_VCF} --haploid --TajimaD 10000  --out ${OUTFILE}

#2
INPUT_VCF="post.s2.recode.vcf"
OUTFILE="TJ_s2_post"

vcftools --vcf ${INPUT_VCF} --haploid --TajimaD 10000 --out ${OUTFILE}

#3
INPUT_VCF="post.s3.recode.vcf"
OUTFILE="TJ_s3_post"

vcftools --vcf ${INPUT_VCF} --haploid --TajimaD 10000  --out ${OUTFILE}

#4
INPUT_VCF="post.s4.recode.vcf"
OUTFILE="TJ_s4_post"

vcftools --vcf ${INPUT_VCF} --haploid --TajimaD 10000  --out ${OUTFILE}

#5
INPUT_VCF="post.s5.recode.vcf"
OUTFILE="TJ_s5_post"

vcftools --vcf ${INPUT_VCF} --haploid --TajimaD 10000 --out ${OUTFILE}
```

Read into R
```{r}
td1 <- read.delim("./Bias_Check/TJ_s1_post.Tajima.D", header = T)
td2 <- read.delim("./Bias_Check/TJ_s2_post.Tajima.D", header = T)
td3 <- read.delim("./Bias_Check/TJ_s3_post.Tajima.D", header = T)
td4 <- read.delim("./Bias_Check/TJ_s4_post.Tajima.D", header = T)
td5 <- read.delim("./Bias_Check/TJ_s5_post.Tajima.D", header = T)
```

```{r}

tdpl1 <- ggplot(td1, aes(x=TajimaD))+
  geom_histogram(bins=20, fill="burlywood3", color="darkslategrey")+
  ggtitle("Sample 1")+
  xlab("")+
  ylab("Tajima's D")

tdpl2 <- ggplot(td2, aes(x=TajimaD))+
  geom_histogram(bins=20, fill="burlywood3", color="darkslategrey")+
  ggtitle("Sample 2")+
  xlab("")+
  ylab("Tajima's D")

tdpl3 <- ggplot(td3, aes(x=TajimaD))+
  geom_histogram(bins=20, fill="burlywood3", color="darkslategrey")+
  ggtitle("Sample 3")+
  xlab("")+
  ylab("Tajima's D")

tdpl4 <- ggplot(td4, aes(x=TajimaD))+
  geom_histogram(bins=20, fill="burlywood3", color="darkslategrey")+
  ggtitle("Sample 4")+
  xlab("")+
  ylab("Tajima's D")

tdpl5 <- ggplot(td5, aes(x=TajimaD))+
  geom_histogram(bins=20, fill="burlywood3", color="darkslategrey")+
  ggtitle("Sample 4")+
  xlab("")+
  ylab("Tajima's D")
```

```{r}
library(gridExtra)
```

```{r}
all_plots <- grid.arrange(pi_pl1, pi_pl2, pi_pl3, pi_pl4, pi_pl5, tdpl1, tdpl2, tdpl3, tdpl4, tdpl5, ncol=2)
ggsave(all_plots, filename = "./Bias_Check/ND_TJ_plots.png", width = 15, height = 30, units = "cm")
```

Comparing this image to this one: /IMGS/diversity_plots.png
They seem to be true to the larger sample of 300-odd, and look very different to the pre-pops.
- tbh this satisfies me, but to be thorough, I should perhaps do the iHS and/or FST

FST is easier
```{bash}
#1
POSTPOP=s1.keep
OUTFILE=s1vspre

vcftools --vcf drought_pops.recode.vcf --haploid --weir-fst-pop preDsamples.txt --weir-fst-pop ${POSTPOP} --out ${OUTFILE}

#2
POSTPOP=s2.keep
OUTFILE=s2vspre

#3
POSTPOP=s3.keep
OUTFILE=s3vspre

#4
POSTPOP=s4.keep
OUTFILE=s4vspre

#5
POSTPOP=s5.keep
OUTFILE=s5vspre
```

I would expect to see really high (>0.9) values in chrs 5, 6, 7, 16
```{r}
fst1 <- read.delim("./Bias_Check/s1vspre.weir.fst", header = T)
high1 <- na.omit(fst1[fst1$WEIR_AND_COCKERHAM_FST>=0.9,])

high1$CHROM <- as.numeric(high1$CHROM)
str(high1)
hist(high1$CHROM)
```

```{r}
fst2 <- read.csv("./Bias_Check/s2vspre.weir.fst", header = T)
high2 <- na.omit(fst2[fst2$WEIR_AND_COCKERHAM_FST>=0.9,])

high2$CHROM <- as.numeric(high2$CHROM)
hist(high2$CHROM)
```

```{r}
fst3 <- read.csv("./Bias_Check/s3vspre.weir.fst", header = T)
high3 <- na.omit(fst3[fst3$WEIR_AND_COCKERHAM_FST>=0.9,])

high3$CHROM <- as.numeric(high3$CHROM)
hist(high3$CHROM)
```
