---
title: "pcadapt"
author: "Georgina Robertson"
date: "19/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Sort out isolate labelling and create all necessary files for downstream analysis.

1) Delete samples WAI327  WAI328 from sample_names and use updated list as keepfile to recode vcf
```{bash}
BASE_VCF="EXCL_HAMIL_CLUS.biallelicSNP.maf0.5.miss0.9"
OUT_VCF="drought_pops"
KEEP_FILE="sample_names.txt"

vcftools --vcf ${BASE_VCF}.recode.vcf --keep ${KEEP_FILE} --recode --out ${OUT_VCF}
```
> Keeping individuals in 'keep' list
After filtering, kept 349 out of 350 Individuals


2) Make sure it worked:
```{bash}
vcf-query -l drought_pops.recode.vcf > sample_names_check.txt
```
Yep

3) Made changes to the assignment of some unknown samples to appropriate drought-pop - edited the source files directly in text editor.

Our populations.
Read in the relevant data:
```{r}
# pre-drought subset
pre <- read.csv("data_sheets/preDsamples.txt", sep = "", header = F)
#28
# post drought subset
post <- read.csv("data_sheets/postDsamples.txt", sep = "", header = F)
#321
# total=349

# this is all the isolates in the VCF
all_in_vcf <- read.csv("FST/sample_names_check.txt", sep = "", header = F)
# total=349
```
Good

4) Need to get exact order of samples in vcf file and get the index value for each member of the two drought-pops

5a) Get the index of the pre-drought samples
```{r}
pre_idx <- match(pre$V1,all_in_vcf$V1)
pre_idx
```
5b) Get the index of the post drought samples
```{r}
post_idx <- match(post$V1,all_in_vcf$V1)
post_idx
```

6) Create drought status identifier for each pop.
Turn each into a dataframe.
Assign column names
```{r}
pre_idx 
pop <- rep("pre", length(pre_idx))
preD <- cbind(pre_idx, pop)
preD_df <- data.frame(preD)
colnames(preD_df) <- c("VCFindex", "popID")

post_idx
pop <- rep("post", length(post_idx))
postD <- cbind(post_idx, pop)
postD_df <- data.frame(postD)
colnames(postD_df) <- c("VCFindex", "popID")
```
7) Combine both pops into one dataframe
Change index value to numeric for...
Sorting by index value (ascending)
```{r}
allD <- rbind(preD_df, postD_df)
allD


allD[, 1] <- lapply(1, function(x) as.numeric(allD[[x]]))
str(allD)

pop_order_index <- allD[order(allD$VCFindex),]
head(pop_order_index)

popIDs_ordered <- pop_order_index$popID
popIDs_ordered
```

## bed file required for pcadpat package 

Make bed file from vcf
```{bash}
plink2 --vcf drought_pops.recode.vcf --make-bed --allow-extra-chr --out drought_pops_pcadapt
## WORKED
```
```{r}
rm(pre)
rm(post)
rm(pre_idx)
rm(post_idx)
rm(preD)
rm(preD_df)
rm(postD)
rm(postD_df)
```


## Proceed to pcadapt_workflow.rmd
