---
title: "Merging SNP index and FST scores"
author: "Georgina Robertson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matching loci from FST output to SNP index as seen in heatmaps

Read in the individual loci files generated by the snp matrices (which were used to produce heatmaps) and merge them into one
```{r}
# set the path and pattern of the files 
files <- list.files(path ="./Processed_SNP_matrices/", pattern = "*loci.csv", full.names = T, recursive = T)

# read them all in while also setting colnames
dflist <- lapply(files, function(i){
 df <- read.csv(i, col.names = c("SNPindex", "Scaffold", "Position"))
 return(df)
})

# rbind them all together
snpindeces <- do.call(rbind, dflist)
```

read in the calculated fst data
```{r}
fst_file <- read.csv("./data_sheets/fst_file.csv")
fst_file <- fst_file[,-1]
```

merge the fst output and the above table by the common values: scaffold and Loci / position
```{r}
FST_SnpIndex <- merge(snpindeces, fst_file, by=c("Position", "Scaffold"))
```

this data is a mess, so will need to sort it
```{r}
FST_SnpIndex <- FST_SnpIndex[order(FST_SnpIndex[,4], FST_SnpIndex[,1]),]
```

Make some unique rownames
```{r}
test <- paste(FST_SnpIndex$Chrom, FST_SnpIndex$Position, sep = ":")

rownames(FST_SnpIndex) <- test
head(FST_SnpIndex)
```


Save this to file!
```{r}
write.csv(FST_SnpIndex ,file="./data_sheets/FST_SNPindex.csv")
```

