---
title: "XP_EHH_Results"
author: "G.Robertson"
date: "14/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Detecting any evidence for a selective sweep in the XP_EHH results


Read in the data
```{r}
XPEHH <- read.csv("./data_sheets/XPEHH_na.rm.csv")
```

Check out the distribution of values
```{r}
ggplot(XPEHH, aes(x=XPEHH_pre_post)) + geom_density()
```
# 99.5% 
```{r}
quantile(XPEHH$XPEHH_pre_post, probs = 0.995)
#    99.5% 
# 2.710898 
```

```{r}
XPEHH99.5 <- XPEHH[XPEHH$XPEHH_pre_post>=2.710898,]
unique(XPEHH99.5$CHR)
# [1] 10 11 12 13  1  2  3  4  5  6  7  8  9
```


# 99.9%
```{r}
quantile(XPEHH$XPEHH_pre_post, probs = 0.999)
#   99.9% 
# 3.35453 
```

```{r}
XPEHH99.9 <- XPEHH[XPEHH$XPEHH_pre_post>=3.35453,]
unique(XPEHH99.9$CHR)
#  [1] 12 13  1  2  3  4  6  7  8  9
```


# 99.95%
```{r}
quantile(XPEHH$XPEHH_pre_post, probs = 0.9995)
#   99.95% 
# 3.699798  
```

```{r}
XPEHH99.95 <- XPEHH[XPEHH$XPEHH_pre_post>=3.699798,]
unique(XPEHH99.95$CHR)
#  [1] 12  1  3  4  8  9
```


# 99.99%
```{r}
quantile(XPEHH$XPEHH_pre_post, probs = 0.9999)
#  99.99% 
# 4.334313 
```

```{r}
XPEHH99.99 <- XPEHH[XPEHH$XPEHH_pre_post>=4.334313,]
unique(XPEHH99.99$CHR)
#  [1] 12  8
```
Write these results. These will be put into an excel sheet.
```{r}
write.csv(XPEHH99.5, file = "./data_sheets/XPEHH_pctl99.5.csv", quote = F, row.names = F)

write.csv(XPEHH99.95, file = "./data_sheets/XPEHH_pctl99.95.csv", quote = F, row.names = F)

write.csv(XPEHH99.99, file = "./data_sheets/XPEHH_pctl99.99.csv", quote = F, row.names = F)
```



Get the top 10% of values for the hline in plots
```{r}
range(XPEHH$XPEHH_pre_post)
# =  -2.699989  4.556530
```

```{r}
2.699989+4.556530
# 7.256519

7.256519/10
# 0.7256519

4.556530 - 0.7256519
# = 3.830878
```
######################## Deprecated #####################
```{r}
highest_XP_EHH <- data[data$XPEHH_pre_post>=2.5, ]
write.csv(highest_XP_EHH, "./data_sheets/highest_XPEHH.csv", row.names = F)
```
Chroms: 2 (7 loci), 8 (1 loci)

```{r}
allhigh <- data[data$XPEHH_pre_post>=2, ]
high_XP_EHH <- a[a$XPEHH_pre_post<2.5,]

write.csv(high_XP_EHH, "./data_sheets/high_XPEHH.csv", row.names = F)

unique(high_XP_EHH$CHR)
```
2720 loci spread across
Chroms: 1  2  3  5  7  8  9 10 11

```{r}
library(ggplot2)
```

A kind of rough way to get a visual representation of where these loci are in thier respective chromosomes.

Pop it out and stretch it to resolve them a bit better. 

```{r}
ggplot(allhigh, aes(x=POSITION/1000, y=CHR))+
  geom_point()+
  xlab("Loci kbp")+
  ylab("Chromosome")+
  ggtitle("Representation of loci with significance from XP-EHH")
```

```{r}
ggplot()+
  geom_point(data = highest_XP_EHH, mapping = aes(x=POSITION, y=CHR),colour="red")+
  geom_point(data = high_XP_EHH, mapping = aes(x=POSITION, y=CHR),colour="blue")+
  xlab("Loci kbp")+
  ylab("Chromosome")+
  ggtitle("Representation of loci with significance from XP-EHH")+
  
```

```{r}
df <- data.frame(x=c(0:129))
ggplot(df, aes(x=x%%10, y=floor(x/10), shape=factor(x), colour=x, size=10)) +
  geom_point() +
  scale_shape_manual(values=df$x) + theme(legend.position='none') +
  scale_x_continuous(breaks=0:10) + scale_y_continuous(breaks=0:12) +
  scale_colour_hue() + scale_colour_gradientn(colours=rainbow(3))
```
start from 0 at bottom left and read right then up (10y + x)
