---
title: "TajD"
author: "G.Robertson"
date: "08/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Boxplot summaries fo TajD and ND per pop per chrom in combined graphs

### Tajima's D

```{r}
taj_pre <- read.csv("./data_sheets/pre_samples.Tajima.D", header = T)

taj_post <- read.csv("./data_sheets/post_samples.Tajima.D", header = T)
```

Extract core chroms, get rid of NAs and change chroms to factors
```{r}
  pre_taj_core <- taj_pre[c(1:3662), ]
  pre_taj_core <- na.omit(pre_taj_core)  
  pre_taj_core$CHROM <- as.factor(pre_taj_core$CHROM)

  post_taj_core <- taj_post[c(1:3668), ]
  post_taj_core <- na.omit(post_taj_core)  
  post_taj_core$CHROM <- as.factor(post_taj_core$CHROM)
```

Making factor variables so they can be coloured by pop
```{r}
a <- taj_post
a$PopID <- rep("post", length(a$CHROM))

b <- taj_pre
b$PopID <- rep("pre", length(b$CHROM))

all_taj <- rbind(a, b)
all_taj_core <- na.omit(subset(all_taj, CHROM<=13))
```

make plot
```{r}
tajD <- ggplot(all_taj_core, aes(x=as.factor(CHROM), y=TajimaD, fill=PopID))+
  geom_boxplot()+
  scale_fill_manual(values = c("burlywood3", "darkslategrey"))+
  xlab("Chromosome")
```

### Nucleotide diversity

```{r}
pi_pre <- read.csv("./data_sheets/pre_samples.windowed.pi", header = T)

pi_post <- read.csv("./data_sheets/post_samples.windowed.pi", header = T)
```

Extract core chroms, get rid of NAs and change chroms to factors
```{r}
 post_pi_core <- pi_post[c(1:3337), ]
  post_pi_core <- na.omit(post_pi_core)  
  post_pi_core$CHROM <- as.factor(post_pi_core$CHROM)

  pre_pi_core <- pi_pre[c(1:3172), ]
  pre_pi_core <- na.omit(pre_pi_core)  
  pre_pi_core$CHROM <- as.factor(pre_pi_core$CHROM)
```


Making factor variables so they can be coloured by pop
Nucleotide diversity
```{r}
a <- post_pi_core
a$PopID <- rep("post", length(a$CHROM))

b <- pre_pi_core
b$PopID <- rep("pre", length(b$CHROM))

all_pi_core <- rbind(a,b)
```

Make plot
```{r}
ndiv <- 
  ggplot(all_pi_core, aes(x=as.factor(CHROM), y=PI, fill=PopID))+
  geom_boxplot()+
  scale_fill_manual(values = c("burlywood3", "darkslategrey"))+
  guides(fill="none")+
  xlab("Chromosome")
  
```

```{r}
library(gridExtra)
```

Put the plots together and save 
```{r}
all_plots <- grid.arrange(ndiv, tajD, ncol=2)
     
 ggsave(all_plots, filename = "./IMGS/ND_TajD_boxplot.png",  width = 30, height = 15, units = "cm")
```

