---
title: "map_plot"
author: "Georgina Robertson"
date: "10/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generating a location map of sample populations

I want a new dataframe for easy plotting of number of isolates per geographical site.

### Inputting and checking the metadata
Read in the excel sheets
```{r}
library("readxl")
packageVersion("readxl")
```
```{r}
all_isolates<- read_excel("data_sheets/WGS_STB_Populations.xlsx", sheet = 1)
summary_pops <- read_excel("data_sheets/WGS_STB_Populations.xlsx", sheet = 2)
isolate_list <- read_excel("data_sheets/WGS_STB_Populations.xlsx", sheet = 3)
```

What variables are available?
```{r}
colnames(all_isolates)
```

How any isolates are in each population?
```{r}
location_count <- as.data.frame(table(all_isolates$Population_name))
# the dimensions match the summary_pops" df
```


###  Creating new dataframe fro easy mapping.

Subsetting some data into a table for map graphing
```{r}
library(dplyr)
```

```{r}
# creating subset
test <- select(all_isolates, COL_YEAR,"Specific Latitude", "Specific Longitude", Population_name, Locality)

# renaming cols
colnames(test) <- c("Year", "Latitude", "Longitude", "Population_name", "Location")

# aggregating duplicatess and getting count of isolates per pop and adding as variable
mapgraphing <- test %>% group_by(Year, Latitude, Longitude, Population_name, Location) %>% summarize((n()))

# renaming new variable
colnames(mapgraphing)[6] <- "PopSize"

# clean up
rm(test)

# writing data to file for future calls
write.csv(mapgraphing, file = "data_sheets/mapgraphing_table.csv")
```

__________________________________________________________________________________
Inputting above subset from file for future reference:
```{r}
mapgraphing <- read.csv("data_sheets/mapgraphing_table.csv", header = T)
mapgraphing
```
__________________________________________________________________________________

Just some urls for map graphing methods:
https://www.r-graph-gallery.com/map.html
https://www.r-graph-gallery.com/bubble-map.html
https://medium.com/analytics-vidhya/mapping-australia-in-r-6ce092c48b49

_______________________________________________________________________________

### Creating the data map

Load the required packages:
```{r}
if(!requireNamespace("devtools")) install.packages("devtools")
devtools::install_github("dkahle/ggmap", ref = "tidyup")
library(ggmap)
register_google(key = "AIzaSyCX8w-wvuSqNRvzF14PFs3XY1S60zTbRdk") 
```
Test to see if the correct map is displayed:
```{r}
p <- ggmap(get_googlemap(center = c(lon = 142.31443124884743, lat = -37.55279466696095), #Mortlake, Vic
                         zoom = 6, scale = 2,
                         maptype ='terrain',
                         color = 'bw'))
p
```

Function to remove all axes text and girdlines etc
```{r}
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
)
```


Set the year as a factor so it can be used in aesthetics
```{r}
mapgraphing$Year <-  as.factor(mapgraphing$Year)
str(mapgraphing)
```

Draw the map with the data:
```{r}
p2 <- p + geom_point(aes(x=Longitude, y=Latitude, size= PopSize, colour = Year),data=mapgraphing,
                     alpha =0.8)+
  ditch_the_axes #run the function above
p2
```

Save the map to file:
```{r}
png(file="IMGS/Pop_size_location.png", width=4800, height=4800, units="px", res=600, pointsize=10)
    # type="cairo", antialias="cleartype")
print(p2)
dev.off()
```

Clean up env:
```{r}
rm(p1)
rm(p2)
rm()
```


